### 2011-06-21 standing-patches-paper.tex
/* The Next Paper */ added some to "local fixation" section

\section{Local establishment}

Include stuff from ``comparison to panmixia'' here:

Note that we are not considering the time it takes the mutation to fix locally.  
How long is this time?  Mean, variance.
When would we expect more mutations in this time?
Roughly, if populations are larger than $C/\mu$ for some $C \approx 100$.

Both these results and the panmictic calculations are underestimates.  
If the time to local fixation is enough to change things significantly, 
we are in the realm where the panmictic results apply better.

Graph: sigma against mean number of origins; different curves for spatial, panmictic (flat), and simulation;
different values of theta as different line types?
Mention Monty's empirical speeds paper.

Compare to panmixia: increases degree of parallel mutation if... ?
Check P and H papers and Orr and Betancourt.  (standing variation only)

>>>>

\section{Local establishment and comparison to panmixia}

In the above (and in \citet{ralphcoop2010}) we have assumed that once the mutation appears,
conditional on eventual fixation, it begins to spread spatially at speed $v$ instantly,
effectively neglecting the time it must first spend fixing locally.
In \citet{ralphcoop2010} we addressed this by noting that there would be no change at all in our results 
if all mutations had to wait the same amount of time before fixing locally,
and that this time was short relative to the time it took the wave to spread across the characteristic length;
we then showed via simulation that this was reasonable in certain situations.
In this section we examine this assumption in more detail, although mostly through heuristic arguments,
and also compare our results above to those of \citet{softsweeps}.

We are assuming that shortly after a new mutation appears, 
it can be approximated by a branching process growing at rate $s$
until the point that it grows large enough to ``feel'' spatial structure,
at which point it begins to spread as a more--or--less deterministic wave.
Although we are not aware of good analysis of this transition, 
the relevant size when spatial structure becomes important
should be something close to $\sigma^2 \rho$ 
(i.e.\ Wright's ``local effective population size'' \citep{XXX}).
Let $Z_t$ be a continuous-time branching process with $Z_0=1$ and $\E[Z_t] = e^{s_b t}$.
Then we know that there exists a random variable $W$ such that $\lim_{t\to\infty} e^{-s_b t} Z_t = W$ almost surely,
so that if $\tau$ is the time $Z_t$ reaches size $\sigma^2 \rho$, i.e.\ $\tau = \inf\{ t \ge 0: Z_t > \sigma^2 \rho$,
then $\sigma^2 \rho = Z_\tau \approx e^{s_b \tau} W$.
From this we know that $\tau \approx (1/s_b) (\log (\sigma^2 \rho) - \log W)$;
although more detailed information is available (e.g.\ a CLT for $\tau$, \citet{XXX}),
we will stick to the loose interpretation.
\marginnote{check jaegers.  isn't there a CLT for $\tau$?}
As $\sigma^2 \rho$ increases, the mean of $\tau$ increases, but the variance approaches a constant
($\var[\tau] \to (1/s_b) \var[\log W]$).

So, roughly speaking, we need to evaluate the importance of a delay of about $T = (1/s_b) \log (\sigma^2 \rho)$.
New mutations will appear and fix during this time if $4N\mu s_b \ge 1/T$,
i.e.\ if $4 N \mu \ge 1/\log (\sigma^2 \rho)$.
Our model will still be a good approximation, however, 
as long as $T$ is short relative to the time a wave takes to spread between nearby mutational origins.
This can be worked out,
but it is simpler to note that if the converse is true, 
that the process is largely unaffected by spatial structure,
and so the panmictic model is a good approximation for the true process.
Since both our spatial model and the panmictic model underestimate the true degree of parallel adaptation,
it suffices to compare results from the two models to avoid applying the wrong model.

% When would we expect more mutations in this time?
% Roughly, if populations are larger than $C/\mu$ for some $C \approx 100$.
% Compare to panmixia: increases degree of parallel mutation if... ?

\citet{softsweepsII}, it is shown that under a panmictic model with certain assumptions on the parameters,
the number of independent origins due to both standing variation and new mutation seen in a sample of size $n$
has approximately the Ewens distribution with parameters $n$ and $\theta = 4 N \mu$.
As $n$ increases, the total number of types seen grows as $\log n$, 
so the ``mean density of types'' corresponding to our $\lambda_0 + \nu_+$ in their model is infinite.
\marginnote{Graph: sigma against mean number of origins; different curves for spatial, panmictic (flat), and simulation;
different values of theta as different line types?  but how to do this?
Mention Monty's empirical speeds paper.}
We could compare the probability of parallel adaptation under the two models (see eq'n 18 in \citet{softsweepsII}),
but this seems less useful than an assessment of the degree of parallel adaptation.
